DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

In the current version of the manuscript and the code, we simply fitted a few $t$ to match the different $\sigma$. In theory, there should be a connection between the two values providing the graph Laplacian has converged toward the Laplace Beltrami operator. Again, this is left for a futur contribution.

\subsection{}
\begin{mdframed}[style=comment]
A question that I would like to know personally, are you thinking about a future upgrade to non-radial kernels?
\end{mdframed}
We are thinking about it. However, as the graph framework we are using only works with radial filters, it is not a simple question. One simple trick to make the network sensitive to direction is to add a per-channel fully connected layers. We are currently investigating other potential graph techniques that would lead to non-radial kernels.

\subsection{Details}
\begin{mdframed}[style=comment]
Finally, I noticed a few points in the text where there were typos or things which were difficult to follow and I have put a note of them below and tried to correct them as much as I can so that you don't have to put too much effort in - I can't wait to see this paper published. Sorry that the list is so extensive (I'm a bit of a stickler for details and trying to write papers concisely):
\end{mdframed}
Thank you very much for taking the time for making these corrections. We find them very helpful and improving the quality of the paper.
We implemented most of them, see below.

\begin{mdframed}[style=comment]
Convolutional Neural Networks (CNNs) have been proposed as an alternative analysis tool in cosmology thanks to their ability to automatically design relevant statistics to maximise the precision of the parameter estimation, while maintaining robustness to noise [9–16].

- There are many of us in the community who are not overly keen on "precision" of parameter estimation being the target for machine learning. Rather, we would prefer robustness of parameter estimation. It might be worth changing this sentence a bit so as not to upset some of the referenced authors ;)
\end{mdframed}

Thank you for this comment. Precision, understood as the final size of the posterior distribution on the measured parameters (including systematic errors), is the key interest of cosmology analysis.
The robustness to noise is also crucial, as demonstrated in [9]. We expanded the sentence to include this.
% \todo{Assigned: @tomek, @michael}
% \TK{not sure what do to here.. will people get upset? "Precision" is commonly used in cosmology}

### 2019-03-04 rebuttal.tex

\begin{mdframed}[style=comment]
Thank you for submitting your manuscript to Astronomy and Computing. I have received comments from reviewers on your manuscript. Your paper should become acceptable for publication pending suitable minor revision and modification of the article in light of the appended reviewer comments.  In particular, there is a request from a reviewer to have access to the data used in your analysis.

Though this is designated a minor revision, there are quite a few points that need to be addressed before the paper is acceptable for publication.  When resubmitting your manuscript, please carefully consider all issues mentioned in the reviewers' comments, outline every change made point by point, and provide suitable rebuttals for any comments not addressed.
\end{mdframed}

Dear Editor,

Thanks for handling our submission. We have revised our manuscript following many of the reviewer suggestions. We believe that it is now ready for publication. Currently, the dataset can be accessed simply via Zenodo. While this process is not anonymous, we provide in this rebuttal (Comment \#2) a special link for the reviewers to access it is anonymously.

@@

\subsection{}
\begin{mdframed}[style=comment]
One fairly significant concern I have is how convincing the application to other machine learning algorithms is. The authors compare deep sphere to a SVM classifier, in the case of distinguishing from two cosmological models from density maps. This classification is obviously a simplification of what would be done in real cosmological analyses, but is sufficient for the purpose of demonstration here. However, the metrics adopted do not clearly highlight the superiority of DeepSphere. The SVM takes pixel histograms or power spectrum densities (PSDs) of each input map. It is not surprising that DeepSphere outperforms them given that it takes the whole map as input. Histograms and PSDs are very restrictive compressions of the data - dropping a lot of spatial information. An SVM or even a standard fully connected neural network (NN) would perform a lot better. It is unclear how that would compare to DeepSphere. But I would certainly expect DeepSphere to be a lot easier to train and have many fewer parameters, which I think is the key advantage that needs to be highlighted here. This is analogous to comparing a standard convolutional NN to a standard fully-connected NN. With appropriate depth both will extract the same information and perform equally well. But of course the CNN has fewer parameters and is easier to train since its structure naturally extracts spatial information. One wouldn't compare the CNN to a SVM trained on pixel histograms or PSDs, since those would not have access to the same information. It would be more natural to compare the CNN and the NN trained on the same maps, and examine the number of parameters, the time and number of examples needed for training to achieve the same performance, etc. This would truly highlight how the structure of the NN is a key advantage.
\end{mdframed}

We agree that the ML algorithms should be tested on the same input data for a fair comparison.

We performed the comparison with these features at it reflects the two-steps traditional way in classic machine learning: a) designing features b) using a method a ``simple'' classifier. Deep learning changed this this tactic with networks that learn the features directly from the raw data.
The PS and histograms were chosen because they are standard statistics used by the cosmology community for analysing real data.
This is why we chose it as the reference point.
In fact, the single step of moving away from these hand-picked statistics to ML approaches is already interesting for physicists.
It would be indeed interesting to compare the performance of the CNN to other ML-based algorithms, but we belive it is a major undertaking if to be conducted fairly.
Given that the our method gives significant advantage over the traditional statistics used in cosmology, we believe that the current scope of the paper is already interesting for cosmologists.
We would be keen to leave more detailed comparisons of ML-algorithms for future work.

@@

%In practice, those models are limited by the computational power and the dataset size.

This being said, we did try to train an SVM on the raw data, but were unable to obtain over $60\%$ accuracy in the three noiseless cases, which is far behind every other tested model.
Practically, the classifier would either overfit or be over-regularized.
We hence decided not to report the raw-data SVM results in the paper.
This information has now been added to the manuscript.
% \nati{FYI, I did try again to be entirely sure about it.}

@@

\todo{Assigned: @Nati, @Michael, @Tomek}
\nati{Please check if you like what I did in light of the discussion we had.}

We share the opinion of the reviewer when he says that 2D CNNs should probably work as well as DeepSphere, ``by taking in a projected 2D image of the sky and using *loads* of filters to learn about the distortion at different parts of the projected sky.'' Hence we decided to add a new experiment in the paper.

Nevertheless, making a comparison is technically not as trivial as it may seems.
Problematically a radial projection is computationally expensive to the point where, because of the dataset augmentation, it becomes significantly more expensive than the actual training of the network.
Hence, we decided to use the geometrical properties of HEALPix to create a ``special'' projection.
In the HEALPix sampling, the sphere is first separated into $12$ parts that are then subdivided into $4$ recursively.
Eventually, each of these $12$ part resembles a square grid.
As the problem we are solving only involve one of these 12 parts, we can use this natural structure to simply re-assign each HEALPix pixel to one of a 2 dimensional grid of size $N_{sides} \times N_{sides}$.
This operation is computationally very cheap and probably spread quite evenly across the resulting image.
This projection has recently been used for a CNN in cosmology in the very recent work \cite{krachmalnicoff2019convolutional}.

As we believe that this projection is the optimal one to use a traditional CNN for application, we performed the experiment using it.

@@

% \mdeff{Agreed that we should do at least one thing they asked for.
% \begin{itemize}
% 	\item Testing on the 1/12 of the sphere is not what he asked for. On one hand, a global projection would exhibit more distortion. On the other, we never use the whole sphere.
% 	\item Keep in mind that there is no orientation that is consistent for the 12 base faces (i.e., where is up?).
% 	\item Agreed that the performance should be very similar. Both are doing local computations. And locally (3 to 5 pixels across), the sphere is isomorphic to an Euclidean plane.
% 	\item The only difference, in my opinion, would be due to the anisotropic filters the 2D CNN has access to.
% 	\item Not sure we should include this result in the main paper. Hard to justify in the story, especially as a comparison with alternative spherical CNNs is missing. Let's see the results.
% \end{itemize}
% }

\subsection{}

\begin{mdframed}[style=comment]
Similarly, I think it would useful to check how well the DeepSphere CNN does against the FCN when the input data is rotated randomly during training - I imagine that the results should match or probably exceed the FCN results and I think this would be a neat little test which doesn't involve needing extra data, just more augmentation.
\end{mdframed}

% That comparison is not warranted.
% or justified. pointless is maybe too aggressive
We do not believe this comparison is useful because of the following theoretical reason
There are two ways to deal with symmetries one wants to be invariant to: (i) build them into the architecture of the model, or (ii) augment the dataset such that a (more general) model learns them ``brute-force''.
With respect to rotation invariance, the FCN architecture is of the first kind, while the CNN is of the second kind.
Assuming the task is truly invariant to said symmetry,\footnote{As argued in the introduction, cosmological maps are rotation invariant.} the performance of the first kind of models would be exactly the same as without augmentation, while the second kind would only catch up after seeing enough data to learn the invariance.
% That is a waste of resources which is only justified if it's not possible to back the desired invariance in the architecture.
Augmentation is only justified if it's not possible to back the desired invariance in the architecture.
% The reviewer himself makes that point in comment 2.8: ``building a NN to suit your purposes''.
We updated the manuscript to more precisely reflect this argumentation.

### 2019-03-04 rebuttal.tex

- You mention your method as part of the "above methods", but then you go on to say that using graphs does allow you to accelerate the method when data is only on part of the sphere.
\end{mdframed}
Indeed, our method can be applied only on a part or the sphere. We have corrected the text.

\begin{mdframed}[style=comment]
By interwinding graph convolutional layers and recurrent layers [65], they can for example model structured time series such as traffic on road networks [66], or recursively complete matrices for recommendation [67].

- I'm not sure what interwinding is, I guess you mean by making GC recurrent, or do you mean using GC and then RL separately? Either way there should be commas around "for example".
\end{mdframed}
\todo{Assign @michael}

\begin{mdframed}[style=comment]
To prevent the cosmological models to be distinguished by their power spectra alone, the maps have been smoothed with a Gaussian kernel of radius 3 arcmins to remove high frequencies ($\ell> 1000$ ).

- This should be "To prevent the cosmological models from being distinguished by their power spectra alone the maps have been smoothed with a Gaussian kernel of radius 3 arcmins to remove high frequencies ($\ell> 1000$ ).
\end{mdframed}
We have corrected this formulation according to the reviewer suggestion.

### 2019-03-04 rebuttal.tex

\begin{mdframed}[style=comment]
... is a measure of the variation of the eigenvector $\bf{u}_i$ is on the graph defined by the Laplacian L.

 - I don't think the "is" after $\bf{u}_i$ should be there.
\end{mdframed}
This has been corrected.

\begin{mdframed}[style=comment]
Given the convolution kernel $h:\mathbb{R}_+ \to \mathbb{R}$,a signal $f\in\mathbb{R}^{N_{pix}}$ on the graph is filtered as

- You don't mention what $\mathbb{R}_+$ is or why $h$ defines that map.
\end{mdframed}
We are not sure what the reviewer means here. $\mathbb{R}_+$ is the ensemble of the real positive numbers and $h$ is simply a function that maps any positive real numbers to another real number.

\nati{This seems clear to me. Specifying what $\mathbb{R}_+$ is would be a bit too much I believe. @all, is it also for you?}

@@

\begin{mdframed}[style=comment]
Similarly as (2), the convolution of the signal $f$ by a kernel $g$ is the scalar product of $f$ with translated versions $T_ig$ of the kernel $g$.

- I think this should be "As with equation (2)" or "Similarly, as in equation (2)" but I can't tell which. Also, it would be really useful if you referred to equations by "equation (x)" rather than just "(x)" throughout - much easier to read that way.
\end{mdframed}
% \todo{Assign: @michael, @tomek}
% \nati{I do not know what we should do.}
Thank you for your suggestion. Ideally we would replace equation references such as "(1)" with "Equation (1)", but given the large number of equation references, we would prefer to stick with a short-form to avoid lengthening the paper further.

\begin{mdframed}[style=comment]
The 1-neighborhood of a node is the set of nodes that are directly connected to it. The k-neighborhood of a node is the set of nodes that are connected to it through paths of length k.

- I'm a little confused here, I think the neighbourhood is every pixel which is closer than the distance of some pixel. The value of k would then be the length of the set of ordered distances for all pixels up to this pixel. Is that correct? For example, take a 3x3 square (on Euclidean space for simplicity), if the central pixel has coordinates (0, 0) then the 1-neighbourhood would be all pixels at a length of 1 pixel or less, i.e. (-1, 0), (+1, 0), (0, -1) and (0, +1) who are all at the same distance so the set of distances is {1} pixel and the length of the set is 1. The 2-neighbourhood would then be all pixels which were a distance $\sqrt{2}$ pixels or less away, i.e. (-1, 0), (+1, 0), (0, -1), (0, +1), (-1, -1), (-1, +1), (+1, -1), (+1, +1) so the set of distances would be {1, $\sqrt{2}$} pixels and the length of the set is 2. Is this correct? If so k is not a length. If this is not correct, I think I misunderstand something and so this should be updated in the text.
\end{mdframed}
% Here the notion of neighboorhood is defined in the context of graphs. Hence there is no coordinates to compute distances.

\todo{Assigned: @michael}

### 2019-03-04 rebuttal.tex

\begin{mdframed}[style=comment]
The two sets of maps were created using the standard cosmological model with two sets of cosmological parameters.

- Can you add that you are going to describe the details more thoroughly in section 4.1.
\end{mdframed}
We have updated the manuscript accordingly.

\begin{mdframed}[style=comment]
Implementation of full cosmological inference requires, however, many more simulations, building the likelihood function, and several other auxiliary tasks.

- Unless you have a cheap neural network based likelihood-free method ;)
\end{mdframed}
Indeed. We added a word "typically" to reflect that there may be cases like this.

@@

- There is not enough information in this caption. Can you explain that there is some obvious peak structure. I think a colourbar on the gnomonic projection would be useful too.
\end{mdframed}
\todo{Assign: @all}
\nati{The problem is that random filters are picky as well and look very similar. Should we say that? I do not think a colorbar is necessary since there are numbered axes on the top plots. Do you agree?}

We decided not to add a colorbar as scale information is already present in the top plot of the figure.

\begin{mdframed}[style=comment]
The fact that a graph representation of the disretised sphere enables an efficient convolution relaxes the iso-latitude constraint on the design of sampling schemes who aim to enable fast convolutions.

- It should be "discretised" and also "who" should be "which" (or "that", I'm not actually sure myself)
\end{mdframed}
We followed the reviewer suggestion (and we went for ``which''). 
% \todo{Assign: @tomek -> \nati{I corrected it with "which". Do you agree? }}
% \TK{Sounds good to me.}

@@

\begin{mdframed}[style=comment]
In the long term, it might enable researchers to consider sampling schemes with different trade-offs, or remove the need for schemes and interpolation altogether and simply consider the positions at which measures were taken.

- What is "it" here? DeepSphere or graph networks or something else?
\end{mdframed}
"it" refer to "graphs" in general. We have updated the manuscript.

\begin{mdframed}[style=comment]
We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of [33]. That is, however, out of the scope of this contribution.

- What makes you believe this? If you don't want to mention it in this paper, please don't speculate a "belief". I might just be reading the sentence incorrectly, but it makes me a little uncomfortable when there is no evidence backing that up. If you have some evidence then it's worth presenting it.
\end{mdframed}
\todo{Assign: @nati, @michael}
As already mentioned in this rebuttal, our belief comes from the fact that we are working on this. 
Nevertheless, we agree with the reviewer that a scientific article should not contains speculations. 
Hence, we removed that suggestion from the manuscript.

@@

\begin{mdframed}[style=comment]
In this case, the graph setting used through this contribution correspond to assume reflective border conditions.

- It should be "assumed". "In this case, the graph setting used through this contribution correspond to assumed reflective border conditions."
\end{mdframed}
Thank you for pointing it out, we believe that the original intended meaning is better reflected as is now.

@@

\begin{mdframed}[style=comment]
Nevertheless, dealing working with a subpart of the sphere, its border slightly affects the convolution.

- There is a superfluous word (dealing/working). I think this would read better as "Nevertheless, when working with only a part of the sphere, the border slightly affects the convolution."
\end{mdframed}
We have corrected this mistake according to the reviewer suggestion.

\begin{mdframed}[style=comment]
These border effects can, however, be mitigated by padding with zeros a small area around the part of interest.

- Are they really mitigated? Surely there are still effects but you are happy with the level?
\end{mdframed}
\todo{Assign: @nati, @michael}
With the zero padding, the border effects become similar to the ones in traditional CNNs. 
This information has been added to the manuscript.

### 2019-03-04 scnn.tex
details fixes

\section{Graph Fourier modes and spherical harmonics}
\label{sec:comparison_spherical_harmonics}
The first 16 eigenvectors $[\b u_1, \ldots, \b u_{16}]$ of the graph Laplacian $\L$, forming the lower part of the graph Fourier basis $\U$, are shown in \figref{graph_harmonics}.
Let us further observe the spectral properties of our constructed spherical graph laplacian $\L$.
Its eigenvalues, shown in \figref{graph_eigenvalues}, are clearly organized in frequency groups of $2\ell + 1$ orders for each degree $\ell$.
We remind the reader that the amplitude of the Laplacian eigenvalue is proportional to the sum of the variations of its associated eigenvector. Furthermore, all spherical harmonics with the same order $\ell$ have the same variation. Hence, the fact that the Laplacian eigenvalues are grouped in blocks of size $2\ell + 1$ is a hint that the graph eigenvectors approximate the spherical harmonics.
In order to push the comparison one step further, we show the correspondence (scalar product) between the subspaces spanned by the graph Fourier modes and the spherical harmonics in \figref{subspace_harmonics_eigenvectors}.
The block-diagonal structure of the matrix, as opposed to diagonal, results the fact that the phase of the graph eigenvector is not aligned with the one of the spherical harmonics.

While these indications show that the constructed graph Fourier basis approximates well the spherical harmonics, one should not forget that the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an important effect on the graph Fourier modes.
First, we believe that they are responsible for energy leaking outside of the block diagonal in \figref{subspace_harmonics_eigenvectors}.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics.
We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is, however, out of the scope of this contribution.
Third, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e., they will span a small part of the sphere.
Those discrepancies result in a convolution operation that is not exactly equivariant to rotation.
Nevertheless, our experiments suggest that these downsides do not have an important effect on the convolution nor hinder the performance of the neural network.

@@

The graph setting used throughout this contribution corresponds to assuming reflective border conditions.
While that is irrelevant when working on the complete sphere (as it has no border), it slightly affects the convolution operation when only a part of the sphere is considered.
As depicted in \figref{border_effects}, a filter localized near a border (via $h(\L) \b \delta_i$) is no longer isotropic.
These border effects can, however, be mitigated by padding with zeros a small area around the part of interest.
We however do not expect these effects to cause any problem as long as the data samples cover the same area.

>>>>

\section{Graph Fourier modes and spherical harmonics}
\label{sec:comparison_spherical_harmonics}
The first 16 eigenvectors $[\b u_1, \ldots, \b u_{16}]$ of the graph Laplacian $\L$, forming the lower part of the graph Fourier basis $\U$, are shown in \figref{graph_harmonics}.
Let us further observe the spectral properties of our constructed spherical graph laplacian $\L$.
Its eigenvalues, shown in \figref{graph_eigenvalues}, are clearly organized in frequency groups of $2\ell + 1$ orders for each degree $\ell$.
We remind the reader that the amplitude of the Laplacian eigenvalue is proportional to the sum of the variations of its associated eigenvector. Furthermore, all spherical harmonics with the same order $\ell$ have the same variation. Hence, the fact that the Laplacian eigenvalues are grouped in blocks of size $2\ell + 1$ is a hint that the graph eigenvectors approximate the spherical harmonics.
In order to push the comparison one step further, we show the correspondence (scalar product) between the subspaces spanned by the graph Fourier modes and the spherical ha, however, out of the scope of this contribution.rmonics in \figref{subspace_harmonics_eigenvectors}.
The block-diagonal structure of the matrix, as opposed to diagonal, results the fact that the phase of the graph eigenvector is not aligned with the one of the spherical harmonics.

While these indications show that the constructed graph Fourier basis approximates well the spherical harmonics, one should not forget that the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an important effect on the graph Fourier modes.
First, we believe that they are responsible for energy leaking outside of the block diagonal in \figref{subspace_harmonics_eigenvectors}.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics.
% We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is, however, out of the scope of this contribution.
The theoretical study of the convergence is, however, out of the scope of this contribution.
Third, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e., they will span a small part of the sphere.
Those discrepancies result in a convolution operation that is not exactly equivariant to rotation.
Nevertheless, our experiments suggest that these downsides do not have an important effect on the convolution nor hinder the performance of the neural network.

@@

The graph setting used throughout this contribution corresponds to assuming reflective border conditions.
While that is irrelevant when working on the complete sphere (as it has no border), it slightly affects the convolution operation when only a part of the sphere is considered.
As depicted in \figref{border_effects}, a filter localized near a border (via $h(\L) \b \delta_i$) is no longer isotropic.
These border effects can, however, be mitigated by padding with zeros a small area around the part of interest. 
In this case, they are similar to the border effects in traditional CNNs.
We however do not expect these effects to cause any problem as long as the data samples cover the same area.
